---
title: "ass3_wk4"
author: "Chuangfu Xie, 27771539"
output: html_document
---
## M1
### (a) 
```{r}
print('Binomial distribution with n=500, theta=0.001')
for (i in c(0:9)){
  cat('When k=',i,', the exact value p(X=k)=',dbinom(i,500,0.001),'\n')
}
cat('p(X<10)=',pbinom(9,500,0.001))
```

### (b) Using Gaussian as approximation
According to central limit theorem, binomial PMF converging to the normal density as n become larger.  
Hence, we can use Gaussian to yields the approximation:
$$E(X_{i})=p, Var(X_{i})=p(1-p)$$
$$X\sim B(n,p)\xrightarrow []{approximate} X\sim N(np,p(1-p))$$
let:
$$ Z=\frac {X-np}{\sqrt{np(1-p)}}\sim N(0,1)$$
Then:
$$Z=\frac {10-500\times 0.001}{\sqrt{500\times 0.001(1-0.001)}}=\frac {9.5}{\sqrt{0.4995}}=8.793$$
$$P(X<10)\approx \Phi (8.793)=1$$
```{r}
m=500*0.001
s=0.001*(1-0.001)
z=(10-m)/(sqrt(500*s))
cat("mean=n*p=",m,', variance=p(1-p)=',s,'\n')
cat("p(X<10)=",pnorm(10,m,s),"\n")
cat("p(Z)=",pnorm(z))
```

### (c) Use Poisson as approximation
Using Poisson approximation can yield s a good aprroximation when n is large and p is small:
$$P(X=i)={C}_{n}^{i} p^{i}(1-p)^{n-i} \approx e^{-\lambda}\frac{\lambda^{i}}{i!}$$
```{r}
lambda=500*0.001
result=0
for (i in c(0:9)){
  cat('When i=',i,', p(X=k)=',dpois(i,lambda),'\n')
  result= result + dpois(i,lambda)
}
cat('P(x<10)=',result)
```

### (d)
$$p=0.001,1-p=0.999$$
$$P(X=i)={C}_{n}^{i} p^{i}(1-p)^{n-i}$$
$$P(X=0)={C}_{500}^{0}p^{0}(1-p)^{500}\approx1\times 0.606=0.606$$
$$P(X=1)={C}_{500}^{1}p^{1}(1-p)^{499}=500\times6.069\times10^{-4}=0.303$$
$$P(X=2)={C}_{500}^{2}p^{2}(1-p)^{498}\approx1.248\times6.076\times10^{-2}=0.076$$
$$P(X=3)={C}_{500}^{3}p^{3}(1-p)^{497}\approx2.071\times6.082\times10^{-3}=0.0126$$
$$P(X=4)={C}_{500}^{4}p^{4}(1-p)^{496}\approx2.573\times6.088\times10^{-4}=1.566\times10^{-3}$$
$$P(X=5)={C}_{500}^{5}p^{5}(1-p)^{495}\approx2.552\times6.094\times10^{-5}=1.555\times10^{-4}$$
$$P(X=6)={C}_{500}^{6}p^{6}(1-p)^{494}\approx2.106\times6.100\times10^{-6}=1.285\times10^{-5}$$
$$P(X=7)={C}_{500}^{7}p^{7}(1-p)^{493}\approx1.486\times6.106\times10^{-7}=9.074\times10^{-7}$$
$$P(X=8)={C}_{500}^{8}p^{8}(1-p)^{492}\approx9.158\times6.113\times10^{-9}=5.598\times10^{-8}$$
$$P(X=9)={C}_{500}^{9}p^{9}(1-p)^{491}\approx5.006\times\times6.119\times10^{-10}= 3.063\times 10^{-9}$$
$$P(X<10)=\sum_{i=1}^{9}P(X=i)\approx1$$


## M2
$$IQ\sim N(100,15^{2})$$
Let's normalise it, say $Z_{IQ}\sim N(0,1)$, then:
$$P(Z_{IQ}>x)= P(\frac{IQ-100}{15}>\frac{x-100}{15})$$
### (1) Probability that 2 or more out of 40 would have IQs greater than 150
$$P(Z_{IQ}\le150)= P(\frac{IQ-100}{15}\le\frac{150-100}{15})=P(Z_{IQ}\le\frac{10}{3})=\Phi(\frac{10}{3})=0.99957$$
$$P(Z_{IQ}\gt150)=1-P(Z_{IQ}\le150)=0.043\%$$
### (2) Probability that 2 or more out of 40 would have IQs greater than 140
$$P(Z_{IQ}\le140)= P(\frac{IQ-100}{15}\le\frac{140-100}{15})=P(Z_{IQ}\le\frac{8}{3})=\Phi(\frac{8}{3})=0.99617$$
$$P(Z_{IQ}\gt140)=1-P(Z_{IQ}\le140)=0.383\%$$